Tulai

### 2023-09-21 paper/paper_tulai.qmd
added grouped_level column. edited depths and chronological units.

  separate(depth, 
           into = c("upper",
                    "lower"),
           sep = "-",
           remove = FALSE,
           convert = TRUE) %>% 
  mutate(lower = parse_number(lower)) %>% 
  rowwise() %>% 
  mutate(midpoint = mean(c(upper, lower))) %>%
# create 'layer' column
  mutate(layer = case_when(
      area == "tp1" & depth == "0-10 cm" ~ "c",
      area == "tp1" & depth == "0-20 cm" ~ "c",
      area == "tp1" & depth == "10-30 cm" ~ "c",
      area == "tp1" & depth == "10-90 cm" ~ "c",
      area == "tp1" & depth == "20-30 cm" ~ "c",
      area == "tp1" & depth == "30-40 cm" ~ "c",
      area == "tp1" & depth == "40-50 cm" ~ "c",
      area == "tp1" & depth == "50-100 cm" ~ "c",
      area == "tp1" & depth == "50-60 cm" ~ "c",
      area == "tp1" & depth == "50-70 cm" ~ "c",
      area == "tp1" & depth == "80-90 cm" ~ "c",
      area == "tp1" & depth == "90-100 cm" ~ "c",
      area == "tp1" & depth == "100-110 cm" ~ "c",
      area == "tp1" & depth == "110-120 cm" ~ "c",
      area == "tp1" & depth == "120-130 cm" ~ "c",
      area == "tp1" & depth == "120-140 cm" ~ "c",
      area == "d1" & depth == "0-20 cm" ~ "c",
      area == "d1" & depth == "20-30 cm" ~ "c",
      area == "d1" & depth == "40-60 cm" ~ "a",
      area == "d1" & depth == "50-60 cm" ~ "a",
      area == "d1" & depth == "60-70 cm" ~ "a",
      area == "d1" & depth == "70-80 cm" ~ "a",
      area == "ta" & depth == "0-30 cm" ~ "e",
      area == "ta" & depth == "30-40 cm" ~ "e",
      area == "ta" & depth == "40-50 cm" ~ "e",
      area == "ta" & depth == "50-60 cm" ~ "e",
      area == "ta" & depth == "60-70 cm" ~ "e",
      area == "ta" & depth == "70-120 cm" ~ "d",
      area == "ta" & depth == "120-150 cm" ~ "d",
      area == "ta" & depth == "150-180 cm" ~ "d",
      area == "ta" & depth == "180-190 cm" ~ "d",
      area == "ta" & depth == "190-200 cm" ~ "d",
      area == "ta" & depth == "200-210 cm" ~ "d",
      area == "ta" & depth == "210-220 cm" ~ "d",
      area == "ta" & depth == "220-230 cm" ~ "d",
      area == "ta" & depth == "230-240 cm" ~ "d",
      area == "ta" & depth == "230-250 cm" ~ "d",
      area == "ta" & depth == "250-260 cm" ~ "d",
      area == "ta" & depth == "260-270 cm" ~ "d",

      TRUE ~ NA_character_
  ))

@@

p2 <- 
tl1 %>% 
  group_by(layer) %>% 
  tally() %>% 
  drop_na(layer) %>% 
ggplot() + 
  aes(layer, n) +
  geom_col()

@@

tl3 <- tl3 %>% 
  filter(!lithic_id %in% c("3154", "1741", "1680"))
```

@@

tl_final <- tl4 %>%
  rename(
    blank =                Blank2,
   `sickle shine` =       shine2,
    length =              `Length (mm)` ,
    width =               `Width (mm)`,
    thickness =           `Thickness (mm)`, 
    typology =             Typology,
    utilization =          Utilization,
    retouch =              Retouch,
    blank =                Blank2, 
   `blank part` =         `Blank_part`,
   `bulb thickness` =     `Bulb Thickness (mm)`, 
   `platform thickness` = `Platform Thickness (mm)`, 
   `platform length` =    `platform Length (mm)`, 
    eraillure =            Eraillure , 
    cortex =               `Cortex (%)`)  %>%
  mutate(al = paste0(toupper(area), "_", layer)) %>%
  relocate(al, .after = layer)   %>%
  select(lithic_id, 
         TA, 
         area, 
         layer,
         al,
         cortex,
         depth, upper, lower, midpoint, 
        `length`, `width`, `thickness`, 
         typology, 
         utilization, 
        `sickle shine`,
         retouch, 
        `typology-tool-core`,
        `tool-typology`,
         blank, `blank part`, 
        `core-typology`, `core-fragment`,`core-technology`, 
        cortex)  %>% 
# Update the 'typology-tool-core' column based on 'Retouch'
  mutate(
    `typology-tool-core` = case_when(
      retouch == 1 ~ "tool",           
      utilization == 1 ~ "tool",
      `sickle shine` == 1 ~ "tool",
      TRUE ~ `typology-tool-core`      
    )
  ) %>% 
  mutate(
    `tool-typology` = case_when(
      utilization == 1 ~ "utilized tool",
      `sickle shine` == 1 ~ "sickle shine",
      TRUE ~ `tool-typology`               
    )
  ) %>% 
  mutate(
    `tool-typology` = case_when(
     # If Retouch is 1 and tool-typology is NA
       retouch == 1 & is.na(`tool-typology`) ~ "retouched piece",
       `sickle shine` == 1 & is.na(`tool-typology`) ~ "sickle shine",

@@

### Summary tables: LAYER

```{r}
# Create a summary table
summary_table_layer <- tl_final %>%
  group_by(layer) %>% 
  summarise(
    `Cores (n)` = sum(!is.na(core)),  # Count non-NA core entries
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  # Count non-NA corefrag entries
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

@@

```{r}
# Generate the table
kable(summary_table_layer, 
      caption = "Summary of lithics by Layer", 
      align = 'c')

@@

### Summary tables: AREA-LAYER:

```{r}
# Create a summary table based on the 'al' column
summary_table_al <- tl_final %>%
  group_by(al) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

@@

kable(summary_table_al, 
      caption = "Summary of lithics by Area and layer")

@@

summary_table_area <- summary_table_area %>%
  add_row(
    area = "Total",
    blade_count =              sum(.$blade_count, na.rm = TRUE),
    flake_count =              sum(.$flake_count, na.rm = TRUE),
    bladelet_count =           sum(.$bladelet_count, na.rm = TRUE),
    tool_count =               sum(.$tool_count, na.rm = TRUE),
    core_count =               sum(.$core_count, na.rm = TRUE),
    core_rejuvenation_count =  sum(.$core_rejuvenation_count, na.rm = TRUE),
    retouched_tool_count =     sum(.$retouched_tool_count, na.rm = TRUE),
    utilized_tool_count =      sum(.$utilized_tool_count, na.rm = TRUE),
    shine_tool_count =         sum(.$shine_tool_count, na.rm = TRUE),
    total =                    sum(.$total, na.rm = TRUE)  # Explicitly setting the total
  )

@@

```{r}
# Prepare data for Summary table-area_layer(al)
summary_table_al <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1")) %>% 
  group_by(al) %>% 
  summarise(
    blade_count = sum(blank == "blade", na.rm = TRUE),
    flake_count = sum(blank == "flake", na.rm = TRUE),
    bladelet_count = sum(blank == "bladelet", na.rm = TRUE),
    tool_count = sum(retouch == 1, na.rm = TRUE) + sum(utilization == 1, na.rm = TRUE),
    core_count = sum(!is.na(`core-typology`), na.rm = TRUE),
    core_rejuvenation_count = sum(!is.na(`corefrag`), na.rm = TRUE),
    retouched_tool_count = sum(retouch == 1, na.rm = TRUE),
    utilized_tool_count = sum(utilization == 1, na.rm = TRUE),
    shine_tool_count = sum(`sickle shine` == 1, na.rm = TRUE),
    total = n()
  )

# Adding a new row with total counts
summary_table_al <- summary_table_al %>% 
  add_row(
    al = "Total",
    blade_count =              sum(.$blade_count, na.rm = TRUE),
    flake_count =              sum(.$flake_count, na.rm = TRUE),
    bladelet_count =           sum(.$bladelet_count, na.rm = TRUE),
    tool_count =               sum(.$tool_count, na.rm = TRUE),
    core_count =               sum(.$core_count, na.rm = TRUE),
    core_rejuvenation_count =  sum(.$core_rejuvenation_count, na.rm = TRUE),
    retouched_tool_count =     sum(.$retouched_tool_count, na.rm = TRUE),
    utilized_tool_count =      sum(.$utilized_tool_count, na.rm = TRUE),
    shine_tool_count =         sum(.$shine_tool_count, na.rm = TRUE),
    total =                    sum(.$total, na.rm = TRUE)  
  )

@@

summary_table_al <- summary_table_al %>%
  filter(!(al %in% c("D1_NA", "TP1_NA")))

kable(
  summary_table_al,
  col.names = c(
    "Area-Layer",
    "Blade",
    "Flake",
    "Bladelet",
    "Tools",
    "Cores",
    "Core Rejuvenations",
    "Retouched Tools",
    "Utilized Tools",
    "Sickle Shine",
    "Total"
  ),
  caption = "Summary of lithics by techno-typological classification")

@@

```{r}
# Create the summary table based on layer
summary_table1_layer <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1")) %>%
  group_by(layer) %>%
  summarise(
    numb_retouched = sum(retouch == 1, na.rm = TRUE),
    numb_cores = sum(!is.na(`core-typology`), na.rm = TRUE),
    total_lithics = n(),
    percent_retouch = round((numb_retouched / total_lithics) * 100, 2)
  )

@@

```{r}
# Create a summary table based on the 'depth', 'area', and 'layer' columns
tl_final <- tl_final %>%
  mutate(depth = gsub("cm", "", depth))

summary_table_depth_area_layer <- tl_final %>%
  group_by(TA, depth, area, layer, upper, lower) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

summary_table_depth_area_layer_to_join <- 
  summary_table_depth_area_layer

# Generate the table
kable(summary_table_depth_area_layer, 
      caption = "Summary of lithics by Depth, Area, and Layer")
```

@@

# Create a full summary table by joining it back to the original summary table based on 'depth' and 'area'
summary_table_depth_area_layer <- summary_table_depth_area_layer %>%
  left_join(summary_table_depth_tool_type, by = c("area", "depth"))

# Remove the existing 'tool_type' if it exists
if ("tool_type" %in% names(summary_table_depth_area_layer)) {
  summary_table_depth_area_layer <- summary_table_depth_area_layer %>%
    select(-tool_type)
}

# Join again based on 'depth', this time being specific about what we want
summary_table_depth_area_layer <- 
  summary_table_depth_area_layer %>%
  left_join(summary_table_depth_tool_type %>% select(depth, tool_type), 
            by = "depth")

@@

kable(summary_table_depth_area_layer, 
      caption = "Summary of lithics by Depth, Area, and Layer")

@@

```{r}
# Specify the exact columns want to keep
desired_columns <-
  c(
    "TA",
    "depth",
    "area",
    "layer",
    "upper",
    "lower",
    "Cores (n)",
    "Core rejuvenations pieces (n)",
    "Blade (n)",
    "Flake (n)",
    "Bladelet (n)",
    "Retouched Tools (n)",
    "Utilized (n)",
    "Sickle Shine (n)",
    "tool_type"
  )  # Only keep this 'tool_type'

# Filter the tl_final dataframe
summary_table_depth_area_layer <- summary_table_depth_area_layer %>%
  filter(area %in% c('ta', 'd1', 'tp1'))

# Generate the table
kable(summary_table_depth_area_layer, 
      caption = "Summary of lithics by Depth, Area, and Layer")
```

@@

# google sheet has 42 rows, and cols for 
# area, layer, level, depth. What is the equivalent above? tp1_c_3_a

@@

# join with lithic data that 
summary_table_depth_area_layer_to_join %>% 
  left_join(depths_divided_tbl) %>% 
  mutate(label = paste(area, 
                       layer, 
                       # level, # need this
                       division_id,
                       sep = "_")) %>% View

@@

# Here, I've imported the table 'Summary by Depth, Area, Layer' into Google Sheets for easier management.
#I thought this format would simplify the process of those depths needed to be splitting up.

@@

# Create the 'label' column
sum_depth_sheet <- sum_depth_sheet %>%
  mutate(label = paste(area, layer, level, sep = "_"))

@@

>>>>

  separate(depth, 
           into = c("upper",
                    "lower"),
           sep = "-",
           remove = FALSE,
           convert = TRUE) %>% 
  mutate(lower = parse_number(lower)) %>% 
  rowwise() %>% 
  mutate(midpoint = mean(c(upper, lower))) %>%
# create 'chronological unit' column
  mutate(chronological_unit = case_when(
      area == "tp1" & depth == "0-10 cm" ~ "c",
      area == "tp1" & depth == "0-20 cm" ~ "c",
      area == "tp1" & depth == "10-30 cm" ~ "c",
      area == "tp1" & depth == "10-90 cm" ~ "c",
      area == "tp1" & depth == "20-30 cm" ~ "c",
      area == "tp1" & depth == "30-40 cm" ~ "c",
      area == "tp1" & depth == "40-50 cm" ~ "c",
      area == "tp1" & depth == "50-100 cm" ~ "c",
      area == "tp1" & depth == "50-60 cm" ~ "b",
      area == "tp1" & depth == "50-70 cm" ~ "b",
      area == "tp1" & depth == "80-90 cm" ~ "b",
      area == "tp1" & depth == "90-100 cm" ~ "b",
      area == "tp1" & depth == "100-110 cm" ~ "b",
      area == "tp1" & depth == "110-120 cm" ~ "b",
      area == "tp1" & depth == "120-130 cm" ~ "b",
      area == "tp1" & depth == "120-140 cm" ~ "b",
      area == "d1" & depth == "0-20 cm" ~ "c",
      area == "d1" & depth == "20-30 cm" ~ "c",
      area == "d1" & depth == "40-60 cm" ~ "a",
      area == "d1" & depth == "50-60 cm" ~ "a",
      area == "d1" & depth == "60-70 cm" ~ "a",
      area == "d1" & depth == "70-80 cm" ~ "a",
      area == "ta" & depth == "0-30 cm" ~ "e",
      area == "ta" & depth == "30-40 cm" ~ "e",
      area == "ta" & depth == "40-50 cm" ~ "e",
      area == "ta" & depth == "50-60 cm" ~ "e",
      area == "ta" & depth == "60-70 cm" ~ "e",
      area == "ta" & depth == "70-120 cm" ~ "d",
      area == "ta" & depth == "120-150 cm" ~ "d",
      area == "ta" & depth == "150-180 cm" ~ "d",
      area == "ta" & depth == "180-190 cm" ~ "d",
      area == "ta" & depth == "190-200 cm" ~ "d",
      area == "ta" & depth == "200-210 cm" ~ "d",
      area == "ta" & depth == "210-220 cm" ~ "d",
      area == "ta" & depth == "220-230 cm" ~ "d",
      area == "ta" & depth == "230-240 cm" ~ "d",
      area == "ta" & depth == "230-250 cm" ~ "d",
      area == "ta" & depth == "250-260 cm" ~ "d",
      area == "ta" & depth == "260-270 cm" ~ "d",

      TRUE ~ NA_character_
  )) %>%

  mutate(
 grouped_level = case_when(
    area == "tp1" & depth == "0-10 cm" ~ "3",
      area == "tp1" & depth == "0-20 cm" ~ "3",
      area == "tp1" & depth == "10-30 cm" ~ "3",
      area == "tp1" & depth == "10-90 cm" ~ "3",
      area == "tp1" & depth == "20-30 cm" ~ "3",
      area == "tp1" & depth == "30-40 cm" ~ "3",
      area == "tp1" & depth == "40-50 cm" ~ "3",
      area == "tp1" & depth == "50-100 cm" ~ "3",
      area == "tp1" & depth == "50-60 cm" ~ "2",
      area == "tp1" & depth == "50-70 cm" ~ "2",
      area == "tp1" & depth == "80-90 cm" ~ "2",
      area == "tp1" & depth == "90-100 cm" ~ "2",
      area == "tp1" & depth == "100-110 cm" ~ "1",
      area == "tp1" & depth == "110-120 cm" ~ "1",
      area == "tp1" & depth == "120-130 cm" ~ "1",
      area == "tp1" & depth == "120-140 cm" ~ "1",
      area == "d1" & depth == "0-20 cm" ~ "3",
      area == "d1" & depth == "20-30 cm" ~ "3",
      area == "d1" & depth == "40-60 cm" ~ "2",
      area == "d1" & depth == "50-60 cm" ~ "2",
      area == "d1" & depth == "60-70 cm" ~ "1",
      area == "d1" & depth == "70-80 cm" ~ "1",
      area == "ta" & depth == "0-30 cm" ~ "6",
      area == "ta" & depth == "30-40 cm" ~ "6",
      area == "ta" & depth == "40-50 cm" ~ "5",
      area == "ta" & depth == "50-60 cm" ~ "5",
      area == "ta" & depth == "60-70 cm" ~ "5",
      area == "ta" & depth == "70-120 cm" ~ "4",
      area == "ta" & depth == "120-150 cm" ~ "4",
      area == "ta" & depth == "150-180 cm" ~ "3",
      area == "ta" & depth == "180-190 cm" ~ "3",
      area == "ta" & depth == "190-200 cm" ~ "2",
      area == "ta" & depth == "200-210 cm" ~ "2",
      area == "ta" & depth == "210-220 cm" ~ "2",
      area == "ta" & depth == "220-230 cm" ~ "2",
      area == "ta" & depth == "230-240 cm" ~ "2",
      area == "ta" & depth == "230-250 cm" ~ "1",
      area == "ta" & depth == "250-260 cm" ~ "1",
      area == "ta" & depth == "260-270 cm" ~ "1",

TRUE ~ NA_character_
  )
)

@@

p2 <- 
tl1 %>% 
  group_by(chronological_unit) %>% 
  tally() %>% 
  drop_na(chronological_unit) %>% 
ggplot() + 
  aes(chronological_unit, n) +
  geom_col()

@@

tl3 <- tl3 %>% 
  filter(!lithic_id %in% c("3154", "1741", "1680", "طبیعی"))
```

@@

tl_final <- tl4 %>%
  rename(
    blank =                Blank2,
   `sickle shine` =       shine2,
    length =              `Length (mm)` ,
    width =               `Width (mm)`,
    thickness =           `Thickness (mm)`, 
    typology =             Typology,
    utilization =          Utilization,
    retouch =              Retouch,
    blank =                Blank2, 
   `blank part` =         `Blank_part`,
   `bulb thickness` =     `Bulb Thickness (mm)`, 
   `platform thickness` = `Platform Thickness (mm)`, 
   `platform length` =    `platform Length (mm)`, 
    eraillure =            Eraillure , 
    cortex =               `Cortex (%)`)  %>%
  mutate(area_unit = paste0(toupper(area), "_", chronological_unit)) %>%
  relocate(area_unit, .after = chronological_unit)   %>%
  relocate(grouped_level, .after = area_unit)   %>%
  select(lithic_id, 
         TA, 
         area, 
         chronological_unit,
         area_unit,
         grouped_level,
         cortex,
         depth, upper, lower, midpoint, 
        `length`, `width`, `thickness`, 
         typology, 
         utilization, 
        `sickle shine`,
         retouch, 
        `typology-tool-core`,
        `tool-typology`,
         blank, `blank part`, 
        `core-typology`, `core-fragment`,`core-technology`, 
        cortex)  %>% 
# Update the 'typology-tool-core' column based on 'Retouch'
  mutate(
    `typology-tool-core` = case_when(
      retouch == 1 ~ "tool",           
      utilization == 1 ~ "tool",
      `sickle shine` == 1 ~ "tool",
      TRUE ~ `typology-tool-core`      
    )
  ) %>% 
  mutate(
    `tool-typology` = case_when(
      utilization == 1 ~ "utilized tool",
      `sickle shine` == 1 ~ "sickle shine",
      TRUE ~ `tool-typology`               
    )
  ) %>% 
  mutate(
    `tool-typology` = case_when(
     # If Retouch is 1 and tool-typology is NA
       retouch == 1 & is.na(`tool-typology`) ~ "retouched piece",
       `sickle shine` == 1 & is.na(`tool-typology`) ~ "sickle shine",

@@

### Summary tables: chronological unit

```{r}
# Create a summary table
summary_table_unit <- tl_final %>%
  group_by(chronological_unit) %>% 
  summarise(
    `Cores (n)` = sum(!is.na(core)),  # Count non-NA core entries
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  # Count non-NA corefrag entries
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

@@

```{r}
# Generate the table
kable(summary_table_unit, 
      caption = "Summary of lithics by Chronological Units", 
      align = 'c')

@@

### Summary tables: AREA-UNIT:

```{r}
# Create a summary table based on the 'area_unit' column
summary_table_au <- tl_final %>%
  group_by(area_unit) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

@@

kable(summary_table_al, 
      caption = "Summary of lithics by Area and chronological unit")

@@

summary_table_area <- summary_table_area %>%
  add_row(
    area = "Total",
    blade_count =              sum(.$blade_count, na.rm = TRUE),
    flake_count =              sum(.$flake_count, na.rm = TRUE),
    bladelet_count =           sum(.$bladelet_count, na.rm = TRUE),
    tool_count =               sum(.$tool_count, na.rm = TRUE),
    core_count =               sum(.$core_count, na.rm = TRUE),
    core_rejuvenation_count =  sum(.$core_rejuvenation_count, na.rm = TRUE),
    retouched_tool_count =     sum(.$retouched_tool_count, na.rm = TRUE),
    utilized_tool_count =      sum(.$utilized_tool_count, na.rm = TRUE),
    shine_tool_count =         sum(.$shine_tool_count, na.rm = TRUE),

@@

```{r}
# Prepare data for Summary table-area_unit(au)
summary_table_au <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1")) %>% 
  group_by(area_unit) %>% 
  summarise(
    blade_count = sum(blank == "blade", na.rm = TRUE),
    flake_count = sum(blank == "flake", na.rm = TRUE),
    bladelet_count = sum(blank == "bladelet", na.rm = TRUE),
    tool_count = sum(retouch == 1, na.rm = TRUE) + sum(utilization == 1, na.rm = TRUE),
    core_count = sum(!is.na(`core-typology`), na.rm = TRUE),
    core_rejuvenation_count = sum(!is.na(`corefrag`), na.rm = TRUE),
    retouched_tool_count = sum(retouch == 1, na.rm = TRUE),
    utilized_tool_count = sum(utilization == 1, na.rm = TRUE),
    shine_tool_count = sum(`sickle shine` == 1, na.rm = TRUE),
    total = n()
  )

# Adding a new row with total counts
summary_table_al <- summary_table_al %>% 
  add_row(
    area_unit = "Total",
    blade_count =              sum(.$blade_count, na.rm = TRUE),
    flake_count =              sum(.$flake_count, na.rm = TRUE),
    bladelet_count =           sum(.$bladelet_count, na.rm = TRUE),
    tool_count =               sum(.$tool_count, na.rm = TRUE),
    core_count =               sum(.$core_count, na.rm = TRUE),
    core_rejuvenation_count =  sum(.$core_rejuvenation_count, na.rm = TRUE),
    retouched_tool_count =     sum(.$retouched_tool_count, na.rm = TRUE),
    utilized_tool_count =      sum(.$utilized_tool_count, na.rm = TRUE),
    shine_tool_count =         sum(.$shine_tool_count, na.rm = TRUE),
    total =                    sum(.$total, na.rm = TRUE)  
  )

@@

summary_table_al <- summary_table_al %>%
  filter(!(area_unit %in% c("D1_NA", "TP1_NA")))

kable(
  summary_table_al,
  col.names = c(
    "Area-unit",
    "Blade",
    "Flake",
    "Bladelet",
    "Tools",
    "Cores",
    "Core Rejuvenations",
    "Retouched Tools",
    "Utilized Tools",
    "Sickle Shine",
    "Total"
  ),
  caption = "Summary of lithics by techno-typological classification")

@@

```{r}
# Create the summary table based on chronological_unit
summary_table1_unit <- tl_final %>%
  filter(area %in% c("ta", "d1", "tp1")) %>%
  group_by(chronological_unit) %>%
  summarise(
    numb_retouched = sum(retouch == 1, na.rm = TRUE),
    numb_cores = sum(!is.na(`core-typology`), na.rm = TRUE),
    total_lithics = n(),
    percent_retouch = round((numb_retouched / total_lithics) * 100, 2)
  )

@@

```{r}
# Create a summary table based on the 'depth', 'area', and 'chronological unit' columns
tl_final <- tl_final %>%
  mutate(depth = gsub("cm", "", depth))

summary_table_depth_area_unit <- tl_final %>%
  group_by(TA, depth, area, chronological_unit, grouped_level, upper, lower) %>%  
  summarise(
    `Cores (n)` = sum(!is.na(core)),  
    `Core rejuvenations pieces (n)` = sum(!is.na(corefrag)),  
    `Blade (n)` = sum(blank == "blade", na.rm = TRUE),
    `Flake (n)` = sum(blank == "flake", na.rm = TRUE),
    `Bladelet (n)` = sum(blank == "bladelet", na.rm = TRUE),
    `Retouched Tools (n)` = sum(retouch == 1, na.rm = TRUE),
    `Utilized (n)` = sum(utilization == 1, na.rm = TRUE),
    `Sickle Shine (n)` = sum(`sickle shine` == 1, na.rm = TRUE)
  )

summary_table_depth_area_unit_to_join <- 
  summary_table_depth_area_unit

# Generate the table
kable(summary_table_depth_area_unit, 
      caption = "Summary of lithics by Depth, Area, and Chronological Unit")
```

@@

# Create a full summary table by joining it back to the original summary table based on 'depth' and 'area'
summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
  left_join(summary_table_depth_tool_type, by = c("area", "depth"))

# Remove the existing 'tool_type' if it exists
if ("tool_type" %in% names(summary_table_depth_area_unit)) {
  summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
    select(-tool_type)
}

# Join again based on 'depth', this time being specific about what we want
summary_table_depth_area_unit <- 
  summary_table_depth_area_unit %>%
  left_join(summary_table_depth_tool_type %>% select(depth, tool_type), 
            by = "depth")

@@

kable(summary_table_depth_area_unit, 
      caption = "Summary of lithics by Depth, Area, and chronological unit")

@@

```{r}
# Specify the exact columns want to keep
desired_columns <-
  c(
    "TA",
    "depth",
    "area",
    "chronological_unit",
    "grouped_level",
    "upper",
    "lower",
    "Cores (n)",
    "Core rejuvenations pieces (n)",
    "Blade (n)",
    "Flake (n)",
    "Bladelet (n)",
    "Retouched Tools (n)",
    "Utilized (n)",
    "Sickle Shine (n)",
    "tool_type"
  )

# Filter the tl_final dataframe
summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
  filter(area %in% c('ta', 'd1', 'tp1'))

# Generate the table
kable(summary_table_depth_area_unit, 
      caption = "Summary of lithics by Depth, Area, and chronological_unit")
```

@@

# google sheet has 42 rows, and cols for 
# area, chronological_unit, level, depth. What is the equivalent above? tp1_c_3_a

@@

# join with lithic data that 
summary_table_depth_area_unit_to_join %>% 
  left_join(depths_divided_tbl) %>% 
  mutate(label = paste(area, 
                       chronological_unit, 
                       # level, # need this
                       division_id,
                       sep = "_")) %>% View

@@

# Here, I've imported the table 'Summary by Depth, Area, chronological_unit' into Google Sheets for easier management.
#I thought this format would simplify the process of those depths needed to be splitting up.

@@

# Create the 'label' column
sum_depth_sheet <- sum_depth_sheet %>%
  mutate(label = paste(area, chronological_unit, grouped_level, sep = "_"))

@@

#try with summary_table_depth_area_unit
```{r}
# Create the 'label' column
summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
  mutate(label = paste(area, chronological_unit, grouped_level, sep = "_"))

# Add a suffix only to duplicates
summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
  group_by(label) %>%
  mutate(suffix = case_when(
    n() == 1 ~ NA_character_,
    TRUE ~ letters[row_number()]
  )) %>%
  ungroup()

# Append the suffix to 'label'
summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
  mutate(label = if_else(is.na(suffix), label, paste0(label, "_", suffix))) %>%
  select(-suffix)

# Relocate 'label' column after 'level'
summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
  relocate(label, .after = grouped_level)

------------------------------------------
# working up to here. I need to add length, width, and thickness columns
------------------------------------------

# Separate `excavation area` into `length` and `width` and calculate `volume`
summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
  separate(`excavation area`, into = c("length", "width"), sep = "\\*") %>%
  mutate(
    length = as.numeric(length),
    width = as.numeric(width),
    `thickness` = `thickness` / 100  # Convert from cm to m
  ) %>%
  mutate(
    volume = length * width * `thickness`  # Calculate volume in m^3
  )

# Calculate lithic volumetric density 
summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
  mutate(
    `lithic_dens` = round(total_lithics / volume, 2)
  )

# Calculate retouch frequency 
summary_table_depth_area_unit <- summary_table_depth_area_unit %>%
  mutate(
    retouch_freq = round(retouch_tool / volume, 2)  
  )
```
